---
title: "AGO_Ovar229_MO39212_Fut_Anal"
author: "Anh Nguyen Duc"
date: "`r Sys.Date()`"
# output: word_document

output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: true
    to_depth: 5
    collapsed: false
    smooth_scroll: false
    number_sections: true
    theme: united
  word_document: default
header-includes: \usepackage{docmute}
classoption: landscape
---

```{r,echo=F,message=F,warning=F}
rm(list=ls())
# source('ADA5W_OPTIMAL_MATCHING_12SEP2018_29522.R')
# source('ADA5W_TOLERANCE_MATCHING_12SEP2018_29522.R')
require(knitr)
require(haven)
require(magrittr)
require(gmodels)
require(tidyverse)
require(ldbounds)
require(gsDesign)
require(msm)
require(xlsx)
require(rpact)
require(survival)
require(survminer)
require(mvtnorm)
# source("../../../../../../../../Methods/R/survivalSampleSizeMisc.R")
# source('../../../../../../../../Methods/Surv/Non_PH/NPH.R')
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, autodep=FALSE)
# source("https://raw.github.roche.com/adaptR/adaptR-tutorials/master/simpleBoundarySummary.R")
set.seed(21)
```

## Supportive techinical resource

+ https://pages.github.roche.com/successR/successR/index.html

+ https://cran.r-project.org/web/packages/bpp/index.html

+ https://onlinelibrary.wiley.com/doi/full/10.1002/pst.1764 

+ https://www.tandfonline.com/doi/full/10.1080/10543406.2014.972508 

+ https://www.sciencedirect.com/science/article/pii/0197245686900048 

## Plan


+ Consider various futility thresholds on the scale of conditional probability (function post_power in bpp package) for both PFS and OS

    + Using the formula in Lema A.1. of https://www.tandfonline.com/doi/full/10.1080/10543406.2014.972508, 
    
    + whereby $\theta$ is always assumed to be the log hazard ratio in the power calculation.

+ Consider various pairs of ``true'' underlying HR for OS and PFS.

+ Consider various number of OS events (also calculate the associated expected months from FPI to reaching target OS event number).

+ Go through each of these combinations / scenarios and calculate

    + Probability of GO decision (false / true contingent on the underlying pairs of OS/PFS HR). 
    
    + Probability of GO decision & with positive read-out (a subset of the above).
    
    + Expected read-out time (under the current pairs of ``true'' OS/PFS HR).
    
    + To avoid simulation one can first consider again the formula from Lemma A.1. of https://www.tandfonline.com/doi/full/10.1080/10543406.2014.972508, where by
    
        + For both OS and PFS, let $d_{fin}$ be the event number at the final / primary analysis, $d_{int}$ be the event number at the futility analysis, $\theta_{suc}$ be the efficacy boundary on the log hazard ratio scale at the final / primary analysis, $\theta_{int}$ be the observed log HR at the futility analysis and $d_2=d_{fin}-d_{int}$. Finally let $\theta$ be the assumed log HR that the events accumulated after the futility analysis will follow, which is always considered to be the log HR used in power calculation. Further more let $\gamma$ be the futility boundary on the conditional power scale.
        
        + Given these notations and Lemma A.1. formula we have: 
        $$
        \begin{align*}
                        & CP_{\theta}\ge\gamma\\
        \Leftrightarrow & \theta_{int}\le d_{int}^{-1}\left(d_{fin}\theta_{suc}-2\sqrt{d_{fin}}\Phi^{-1}\left(\gamma\right)-d_2\theta\right) 
        \end{align*}
        $$
        
        + For each of the true underlying log HR $\theta_0$ we have 
        
        $$
        \left(
        \begin{array}{c}
        \theta_{int} \\
        \theta_{fin}
        \end{array}
        \right)
        \sim\mathcal{N}
        \left(
        \left(        
        \begin{array}{c}
        \theta_0 \\
        \theta_0 
        \end{array}
        \right),\     
        \left(        
        \begin{array}{cc}        
        \frac{4}{d_{int}} & \frac{4}{d_{fin}}\\ 
        \frac{4}{d_{fin}} & \frac{4}{d_{fin}}
        \end{array}
        \right)        
        \right)        
        $$
        
        + With this one could then calculate the following quantities
        
            + $P_{\theta_0}\left(CP_{\theta}\ge\gamma\right)$
            
            + $P_{\theta_0}\left(CP_{\theta}\ge\gamma\bigcap\theta_{fin}\le\theta_{suc}\right)$
            
            + $P_{\theta_0}\left(\theta_{fin}\le\theta_{suc}\mid CP_{\theta}\ge\gamma\right)$
            
        <!-- + One can further assume independence between PFS and OS, which is likely to be assumed even via simulation, to calculate the probability of having at least one or both positive outcomes. -->
    
+ Caveat of current proposal and possible extensions

    + In the simulation we ignore the intrinsic correlation between PFS and OS as well as the potential to recycle $\alpha$ between the endpoints.
    
    + The use of predicive probability would be mentioned as a potential modification with the caveat concerning the difficulty in choosing a prior.
    
    + An efficacy stopping boundary could also be added.

```{r,echo=T,message=F,warning=F}
## Trial setup -----------------------------
fpi <- as.Date('2018-08-01') 

mOS_ctr <- 15
alp_OS  <- .045 # two-sided
pow_OS  <- 0.8

mPFS_ctr<- 6.5
alp_PFS <- .005 # two-sided
pow_PFS <- 0.92

final_OS_ev <- 391
final_PFS_ev<- 547

rr <- 1 # Pla : Atz

# below data is from file MO39212_Amd1_20210119_V4_0_CI_PST_revRocheLegal_BWJlegal_20210301_20210304_ba_revNC_PST_revNC.doc / Attachment 8 (p22)
tmp_accrualIntensity <-  c(0, 100, 266, 332, 399, 465, 531, 598, 664) %>% diff
tmp_accrualTime <- (((as.Date(c('2018-08-01', '2020-05-01', '2021-04-01', '2021-07-01',
                         '2021-08-01', '2021-10-01', '2021-12-01', '2022-02-01', 
                         '2022-03-01')) - fpi) %>% as.numeric)/30.4375) %>% round
tmp_accrualTime_diff <- tmp_accrualTime %>% diff

accrualIntensity <- accrualTime <- NULL
accrualTime <- 0

for(id in 1:length(tmp_accrualIntensity)) {
  accrualIntensity <- c(accrualIntensity, 
                        rep(tmp_accrualIntensity[id]/tmp_accrualTime_diff[id], times=tmp_accrualTime_diff[id]) )
  accrualTime <- c(accrualTime, (tmp_accrualTime[id]+1):tmp_accrualTime[id+1])
  
} # end of for(id in 1:length(tmp_accrualIntensity))

drop_out_time <- 9 # this would generally only affect timeline

## Simulation / Calculation setup -----------------------------

# true underlying HRs
hrOS  <- c(0.75, 0.8, .95)
hrPFS <- c(0.70, 0.8, .95)

# number of OS events at which the futility analyis is to take place
futi_OS_ev <- c(175, 200, 225, 250)

# futility threshold on (currently) conditional probability scale
futi_threshold <- c(0.2, .25, 0.3, 0.35, .4) 

settings <- expand.grid('hrOS'=hrOS, 'hrPFS'=hrPFS, 'futi_OS_ev'=futi_OS_ev,
                        'futi_threshold'=futi_threshold)
settings <- settings[order(settings$hrOS, settings$hrPFS, settings$futi_OS_ev, settings$futi_threshold),]

deg_OS <- getDesignGroupSequential(kMax=1, alpha=alp_OS/2,  sided = 1, beta=1-pow_OS)

sam_OS <- deg_OS %>% getSampleSizeSurvival(median2 = mOS_ctr, hazardRatio = 0.75, allocationRatioPlanned = rr, accrualTime = accrualTime, accrualIntensity = accrualIntensity, dropoutRate1 = .05, dropoutRate2 = .05, dropoutTime = drop_out_time)

sam_OS %>% summary

deg_PFS<- getDesignGroupSequential(kMax=1, alpha=alp_PFS/2, sided = 1, beta=1-pow_PFS)

sam_PFS<- deg_PFS %>% getSampleSizeSurvival(median2 = mPFS_ctr, hazardRatio = 0.70, allocationRatioPlanned = rr, accrualTime = accrualTime, accrualIntensity = accrualIntensity, dropoutRate1 = .05, dropoutRate2 = .05, dropoutTime = drop_out_time)

theta_suc_OS <- sam_OS$criticalValuesEffectScale %>% as.numeric %>% log
theta_suc_PFS<- sam_PFS$criticalValuesEffectScale %>% as.numeric %>% log

set.seed(21)
out <- NULL

for (id in 1:nrow(settings)) {
  
  # get expected number of PFS events at each of the above OS events 
  tmp <- deg_OS %>% getPowerSurvival(median2 = mOS_ctr, hazardRatio = settings$hrOS[id], allocationRatioPlanned = rr, accrualTime = accrualTime, accrualIntensity = accrualIntensity, dropoutRate1 = .05, dropoutRate2 = .05, dropoutTime = drop_out_time, maxNumberOfSubjects = accrualIntensity %>% sum, maxNumberOfEvents = settings$futi_OS_ev[id])
  
  time_to_OS<- tmp$analysisTime %>% ceiling
  
  OS_calen  <- (fpi + time_to_OS * 30.4375) %>% format("%Y-%b") %>% as.character
  
  event_PFS <- (getEventProbabilities(time = time_to_OS %>% as.numeric, 
                                     accrualTime = accrualTime, accrualIntensity = accrualIntensity,
                                     lambda2 = log(2)/mPFS_ctr, 
                                     hazardRatio = settings$hrPFS[id], allocationRatioPlanned = rr,
                                     dropoutRate1 = .05, dropoutRate2 = .05, dropoutTime = 12)$overallEventProbabilities * sum(accrualIntensity) )%>% ceiling  
  
  d2_OS <- final_OS_ev  - settings$futi_OS_ev[id]
  d2_PFS<- final_PFS_ev - event_PFS
  
  tmp_OS <- (1/settings$futi_OS_ev[id]) * (final_OS_ev * theta_suc_OS - 
                                           2*sqrt(final_OS_ev)*qnorm(settings$futi_threshold[id])-
                                           d2_OS*0.75%>%log)
  
  tmp_PFS <- (1/event_PFS) * (final_PFS_ev * theta_suc_PFS - 
                                           2*sqrt(final_PFS_ev)*qnorm(settings$futi_threshold[id])-
                                           d2_PFS*0.70%>%log)  
  
  prob_cp_ge_futi_threshold_OS <- pnorm(q = tmp_OS, mean = settings$hrOS[id]%>%log, 
                                        sd = 2/sqrt(settings$futi_OS_ev[id]), lower.tail = T)
  
  prob_cp_ge_futi_threshold_PFS<- pnorm(q = tmp_PFS, mean = settings$hrPFS[id]%>%log, 
                                        sd = 2/sqrt(event_PFS), lower.tail = T)  
  
  
  prob_cp_ge_futi_threshold_OS_and_pos_OS <- pmvnorm(lower=c(-Inf, -Inf), 
                                                     upper=c(tmp_OS, theta_suc_OS), 
                                                     mean =rep(settings$hrOS[id]%>%log, 2),
                                                     sigma=rbind(
                                                       c(4/settings$futi_OS_ev[id], 4/final_OS_ev),
                                                       c(4/final_OS_ev,             4/final_OS_ev)
                                                       )
                                                     )
  
  prob_cp_ge_futi_threshold_PFS_and_pos_PFS <- pmvnorm(lower=c(-Inf, -Inf), 
                                                       upper=c(tmp_PFS, theta_suc_PFS), 
                                                       mean =rep(settings$hrPFS[id]%>%log, 2),
                                                       sigma=rbind(
                                                       c(4/event_PFS,    4/final_PFS_ev),
                                                       c(4/final_PFS_ev, 4/final_PFS_ev)
                                                       )
                                                    )  
  
  out <- rbind(out, 
               c('OS_time'=time_to_OS, 'OS_date'=OS_calen, 'PFS_ev'= event_PFS,
                 'P_go_OS'=prob_cp_ge_futi_threshold_OS %>% round(digits=2),
                 'P_go_pos_OS'=prob_cp_ge_futi_threshold_OS_and_pos_OS %>% round(digits=2),
                 'P_pos_OS_cond_go_OS'=(prob_cp_ge_futi_threshold_OS_and_pos_OS/
                                       prob_cp_ge_futi_threshold_OS) %>% round(digits=2),
                 'P_go_PFS'=prob_cp_ge_futi_threshold_PFS %>% round(digits=2),
                 'P_go_pos_PFS'=prob_cp_ge_futi_threshold_PFS_and_pos_PFS %>% round(digits=2),
                 'P_pos_PFS_cond_go_PFS'=(prob_cp_ge_futi_threshold_PFS_and_pos_PFS/
                                       prob_cp_ge_futi_threshold_PFS) %>% round(digits=2)           
                 ))
  
} # end of for (id in 1:nrow(settings))
```

```{r,echo=T,message=F,warning=F,results='asis'}
cbind(settings, out) %>% kable(digits=2) %>% print
out <- out %>% as.data.frame
library(lattice)

for (hr_os in hrOS) {
  
  for (hr_pfs in hrPFS) {  
    tmp_settings <- settings[which(settings$hrOS==hr_os & settings$hrPFS==hr_pfs),]
    tmp_out      <- out[which(settings$hrOS==hr_os & settings$hrPFS==hr_pfs),]
    tmp <- cbind(tmp_settings, tmp_out)
    
    #par(mfrow=c(4,1))    
    
    with(tmp, {
      bp<-barplot(height = futi_OS_ev %>% unique, names.arg = OS_date %>% unique,  col='lightgray', 
                  ylab = 'OS events', main=paste0('True HR OS = ', hr_os, ', HR PFS = ', hr_pfs))
      text(x=bp, y = futi_OS_ev%>%unique/2, labels = futi_OS_ev%>%unique)
      
      bp<-barplot(height = PFS_ev %>% unique %>% as.character %>% as.numeric, 
                  names.arg = OS_date %>% unique,  col='green', ylab = 'PFS events',
                  main=paste0('True HR OS = ', hr_os, ', HR PFS = ', hr_pfs))
      text(x=bp, y = (PFS_ev%>%unique%>%as.character %>% as.numeric)/2, 
           labels = PFS_ev%>%unique%>%as.character %>% as.numeric)      
      
      })
      
    barchart(I(P_go_OS %>% as.character %>% as.numeric)~I(futi_OS_ev%>%as.factor), data=tmp, groups=futi_threshold, 
             scales=list(x=list(rot=45,cex=0.8)), auto.key=T, 
             xlab='OS events at futility analysis', ylab='Prob. of no futility stop for OS', 
             main=paste0('True HR OS = ', hr_os, ', HR PFS = ', hr_pfs)) %>% print        
    
    barchart(I(P_go_PFS %>% as.character %>% as.numeric)~I(PFS_ev), data=tmp, groups=futi_threshold, 
             scales=list(x=list(rot=45,cex=0.8)), auto.key=T, 
             xlab='PFS events at futility analysis', ylab='Prob. of no futility stop for PFS', 
             main=paste0('True HR OS = ', hr_os, ', HR PFS = ', hr_pfs)) %>% print  

  } # end of for (hr_pfs in hrPFS)
  
} # end of for (hr_os in hrOS)

## get results for green, blue and yellow scenario


# out_summary_low <- out_summary_upp <- list()

for (id0 in 1:length(futi_OS_ev)) {
  
  out_summary_low <- out_summary_upp <- matrix(NA, nrow=length(hrOS), ncol=length(futi_threshold)) 
  
  for (id in 1:length(futi_threshold)) {
    
    id_green <- which((settings$hrOS==0.75 | settings$hrPFS==0.70) & settings$futi_threshold==futi_threshold[id] & settings$futi_OS_ev==futi_OS_ev[id0])
    id_blue  <- which(((settings$hrOS==0.80 & settings$hrPFS!=0.70) | 
                      (settings$hrOS!=0.75 & settings$hrPFS==0.80)) & settings$futi_threshold==futi_threshold[id] & settings$futi_OS_ev==futi_OS_ev[id0])
    id_yellow<- which((settings$hrOS==0.95 & settings$hrPFS==0.95) & settings$futi_threshold==futi_threshold[id] & settings$futi_OS_ev==futi_OS_ev[id0])    
    out_green <- out[id_green,]
    tmp_green <- c(pmax(out_green$P_go_OS %>% as.character %>% as.numeric, out_green$P_go_PFS %>% as.character %>% as.numeric),
                   pmin(1, out_green$P_go_OS %>% as.character %>% as.numeric + out_green$P_go_PFS %>% as.character %>% as.numeric)
                  )      
  
    
    out_blue <- out[id_blue,]
    tmp_blue <- c(pmax(out_blue$P_go_OS %>% as.character %>% as.numeric, out_blue$P_go_PFS %>% as.character %>% as.numeric),
                   pmin(1, out_blue$P_go_OS %>% as.character %>% as.numeric + out_blue$P_go_PFS %>% as.character %>% as.numeric)
                  )  
    
    out_yellow <- out[id_yellow,]
    tmp_yellow <- c(pmax(out_yellow$P_go_OS %>% as.character %>% as.numeric, out_yellow$P_go_PFS %>% as.character %>% as.numeric),
                   pmin(1, out_yellow$P_go_OS %>% as.character %>% as.numeric + out_yellow$P_go_PFS %>% as.character %>% as.numeric)
                  )  
    
    
    out_summary_low[1, id] <- min(tmp_green)
    out_summary_upp[1, id] <- max(tmp_green)
    
    out_summary_low[2, id] <- min(tmp_blue)
    out_summary_upp[2, id] <- max(tmp_blue)  
    
    out_summary_low[3, id] <- min(tmp_yellow)
    out_summary_upp[3, id] <- max(tmp_yellow)    
  } # end of for (id in 1:length(futi_threshold))
  out_summary_low %>% kable(col.names = futi_threshold, caption=paste0('Lower bound for Prob of trial continuation, OS ev = ', futi_OS_ev[id0])) %>% print
  out_summary_upp %>% kable(col.names = futi_threshold, caption=paste0('Upper bound for Prob of trial continuation, OS ev = ', futi_OS_ev[id0])) %>% print  
} # end of for (id0 in 1:length(futi_OS_ev))


```

